---
title: "R/Bioconductor tools for mass spectrometry-based proteomics"
author: "[Laurent Gatto](https://github.com/lgatto)"
  BiocStyle::html_document:
     toc: true
     toc_depth: 1
vignette: >
  % \VignetteIndexEntry{R/Bioconductor tools for mass spectrometry-based proteomics}
  % \VignetteEngine{knitr::rmarkdown}
  % \VignetteKeyword{proteomics, mass spectrometry, data}
---


```{r style, echo = FALSE, results = 'asis', message=FALSE}
BiocStyle::markdown()
```

```{r env, echo=FALSE}
suppressPackageStartupMessages(library("RforProteomics"))
```

**Abstract** In this workshop, we will use R/Bioconductor packages to
explore, process, visualise and understand mass spectrometry-based
proteomics data, starting with raw data, and proceeding with
identification and quantitation data, discussing some of their
peculiarities compared to sequencing data along the way. The
participants will gain a general overview of Bioconductor packages for
mass spectrometry and proteomics, and learn how to navigate raw data
and reconstruct quantitative data. The workshop is aimed at a beginner
to intermediate level, such as, for example, seasoned R users who want
to get started with mass spectrometry and proteomics, or proteomics
practitioners who want to familiarise themselves with R and
Bioconductor infrastructure.

This short tutorial is part of the `ProteomicsBioc2016Workshop`
package (version `r packageVersion("ProteomicsBioc2016Workshop")`),
available at https://github.com/lgatto/ProteomicsBioc2016Workshop.


------------
> This vignette available under a
> [**creative common CC-BY**](http://creativecommons.org/licenses/by/4.0/)
> license. You are free to **share** (copy and redistribute the
> material in any medium or format) and **adapt** (remix, transform,
> and build upon the material) for any purpose, even commercially.
------------

# Introduction

```{r, pk, echo=FALSE, warning=FALSE, cache=TRUE}
biocv <- as.character(biocVersion())
pp <- proteomicsPackages(biocv)
msp <- massSpectrometryPackages(biocv)
msdp <- massSpectrometryDataPackages(biocv)
```

In Bioconductor version `r biocv`, there are respectively `r nrow(pp)`
proteomics and `r nrow(msp)` mass spectrometry software packages and
`r nrow(msdp)` mass spectrometry experiment packages. These respective
packages can be extracted with the `proteomicsPackages()`,
`massSpectrometryPackages()` and `massSpectrometryDataPackages()` and
explored interactively.


```{r, pp, eval=FALSE}
library("RforProteomics")
pp <- proteomicsPackages("3.3")
display(pp)
```


# Session information

The source used to generate this document is available [here](https://github.com/lgatto/ProteomicsBioc2016Workshop/blob/master/vignettes/Bioc2016.Rmd).

```{r, si, echo=FALSE}
print(sessionInfo(), local = FALSE)
```
